---
title: "Predicting of the pulmunary TB cases using clinical data"
author: "Patrick Kaggwa"
date: "12/03/2024"
output: html_document
---


# Processing script

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(readr)
library(tidymodels)
```


# Data loading



```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","TB data.xlsx")

#load data
# Loading the needed datasets
madaproject <- read_excel(data_location)

```


# Check data

First we can look at the codebook

```{r}
codebook <- readxl::read_excel(data_location, sheet ="codebook")
print(codebook)
```


Several ways of looking at the data

```{r}
dplyr::glimpse(madaproject)
summary(madaproject)
head(madaproject)
skimr::skim(madaproject)
```


#Cleaning
By inspecting the data as done above, we find some of the data does a missing value


Now I will at the data types

```{r}
#Now I will look at the data types of the variables
str(madaproject)
colnames(madaproject)
```
Now I see that most of the variables are characters yet most of are supposed to be categorical and i see home fuel being numeric yet its sussposed to be categorical 

```{r}

# converting character variables as factors 
madaproject <- mutate_at(madaproject, 
                         vars(Gender, Asthma, Smoke, Alcohol, FeverNO, Weightloss, 
                              CoughNO, `Sputumproduction`, `BloodinSputum`, 
                              cxrconclusion_chestxray, homefuel, HIVrslt, Spatum_Culture), 
                         as.factor)


```
Then I will look and see if the structure has changed 

```{r}
# now I will look at the strusture
str(madaproject)


```
since now the variables have the right data types I will look again to ensure i have no missing values 

```{r}
#generally I will look at missing generally
naniar::gg_miss_var(madaproject)
```
No I see I have some missing data in the outcome variable , now i will list the nas

```{r}
# Check for missing values in each column
na_counts <- colSums(is.na(madaproject))

# Display columns with missing values
na_columns <- names(na_counts[na_counts > 0])
na_columns
```
I see I have one , I will remove it from column BloodinSputum since its one it might not cause an impact
```{r}
madaproject <- na.omit(madaproject)
```

Also I see Home Fuel has 0 and 1 so I will rename the levers to yes = 1 and No = 0

```{r}
# Rename levels
madaproject$homefuel <- ifelse(madaproject$homefuel == "1", "Yes", "No")
```

Testing to comfirm that there is no missing values.
```{r}
#generally I will look at missing generally
naniar::gg_miss_var(madaproject)
```


I will now rename some variables to make them more understable
```{r}
# Renaming the some of the variables 
madaproject <- madaproject %>% 
              rename(tb = Spatum_Culture,
                     feverstatus = FeverNO,
                     coughstatus = CoughNO,
                     hivstatus = HIVrslt)

```

Now I will check my variable
```{r}
colnames(madaproject)
```
Good  the variables changed 

I feel the data is clean I will go ahead and save it has my final dataset

# Save data 
Finally, I will save the clean data as RDS file.


```{r}
# Save finaldata as RDS
save_data_location <- here::here("data","processed-data","madaproject.rds")
saveRDS(madaproject, file = save_data_location)
```







# Notes

If there is any issue with this file and the data created please let me so that I fix the issues.

